# Communicable, Maternal, Neonatal, and Nutritional Disease Burden (CMNN)

```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the communicable diseases data
data_cmnn <- read_csv(here("data", "burden-of-disease-cmnn.csv"))
```

## Table of Estimates for CMNN Burden Over Time

```{r}
# Here render a table for the DALY burden over time for the three countries 
# You should pivot the data to show each country in a separate column.
# Use kable() from the knitr package to render the table.
cmnn_filtered <- data_cmnn %>% 
  filter(Entity %in% c("Egypt", "Uganda", "Nigeria")) %>% 
  select(-Code) %>%
  pivot_wider(names_from = Entity, values_from = `DALYs (Disability-Adjusted Life Years) - Communicable, maternal, neonatal, and nutritional diseases - Sex: Both - Age: Age-standardized (Rate)`) %>% 
  kable()

cmnn_filtered
```

## Plot Showing Trends in CMNN Burden Over Time

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
data_cmnn %>% 
  rename(Country = Entity, `DALYs lost` = `DALYs (Disability-Adjusted Life Years) - Communicable, maternal, neonatal, and nutritional diseases - Sex: Both - Age: Age-standardized (Rate)`) %>% 
  filter(Country %in% c("Egypt", "Uganda", "Nigeria")) %>% 
  ggplot(aes(Year, `DALYs lost`, color = Country)) +
  geom_line()
```

## Summary of CMNN Burden Findings

Nigeria and Uganda have a high CMNN disease burden that both decreases over time.
Egypt has a 3-4 fold lower CMNN disease burden that also decreases over time.
The CMNN disease burden of Nigeria and Uganda decreases at a higher rate than that of Egypt.


